---
title: "NCC: R package for comparing treatment arms against control in platform trials with non-concurrent controls"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      error = FALSE,
                      warning = FALSE,
                      comment = "",
                      fig.align = "center",
                      out.width = "80%")

source("R/data_sim_cont.R")
source("R/linear_trend.R")
source("R/get_ss_matrix.R")
```

## NCC

`NCC` allows users to simulate platform trials and to compare arms using non-concurrent control data. 
This package contains the following functions:

xx

### Overview design 

## Installation

```{r, message=FALSE, warning=FALSE, results=FALSE, eval=FALSE}
# install.packages("devtools") 
devtools::install_github("pavlakrotka/NCC")
```

## Documentation

```{r, message=FALSE, warning=FALSE, results=FALSE, eval=FALSE}
browseVignettes("NCC")
```

## References


---

**Funding**

[EU-PEARL](https://eu-pearl.eu/) (EU Patient-cEntric clinicAl tRial pLatforms) project has received funding from the Innovative Medicines Initiative (IMI) 2 Joint Undertaking (JU) under grant agreement No 853966. This Joint Undertaking receives support from the European Union’s Horizon 2020 research and innovation programme and EFPIA andChildren’s Tumor Foundation, Global Alliance for TB Drug Development non-profit organisation, Spring works Therapeutics Inc. This publication reflects the authors’ views. Neither IMI nor the European Union, EFPIA, or any Associated Partners are responsible for any use that may be made of the information contained herein.

