---
title: "How to install the NCC package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to install the NCC package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE,
               message = FALSE,
               error = FALSE,
               warning = FALSE,
               comment = "",
               fig.align = "center",
               out.width = "70%")
```


Prior to installing the `NCC` package, the [JAGS](https://mcmc-jags.sourceforge.io/) library needs to be installed on your computer, since the Bayesian analysis methods are performed using the [rjags R-package](https://cran.r-project.org/package=rjags) and therefore require the installation of the JAGS library. You can download the latest version of JAGS from [here](https://sourceforge.net/projects/mcmc-jags/files/).

To install the stable version of the `NCC` package from [CRAN](https://cran.r-project.org/package=NCC), please run the following code:

```{r, message=FALSE, warning=FALSE, results=FALSE, eval=FALSE}
install.packages("NCC")
```

To install the latest development version of the `NCC` package from GitHub, please run the following code:

```{r, message=FALSE, warning=FALSE, results=FALSE, eval=FALSE}
# install.packages("devtools") 
devtools::install_github("pavlakrotka/NCC", build_vignettes = TRUE)
```



