---
title: "`data_sim_cont()`: simulating platform trials with continuous data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(NCC)
```


## Short description

The function `data_sim_cont()` enables data simulation of a platform trial with an arbitrary number of treatment arms entering at different time points.


### Assumptions

- sample sizes are equal across all treatment arms
- block randomization is used, a factor to multiply the number of active arms with in order to get the block size in each period can be specified as input argument (period_blocks, default=2)

## Usage options

### Inputs

The user has two options to specify input arguments:

#### 1)

Specify total sample size in the trial (n_total), number of treatment arms (num_arms) and the timing of adding new arms in terms of number of patients allocated to the control arm (t_arm). The sample size per period is then computed using the `get_ss_matrix()` function. The sample sizes in treatment and control arms are varying, based on the 3 specified parameters and the assumptions of equal sample sizes for all treatment arms and 1:1:...:1 allocation ratio in all periods.

```{r, eval=FALSE}
# Example

head(data_sim_cont(n_total = 1000, num_arms = 3, t_arm = 120, 
                   n_arm = NULL, alloc_ratios = NULL, 
                   period_blocks = 2, delta = rep(0.25, 3), lambda = rep(0.15, 4), 
                   sigma = 1, trend = "linear"))
```


#### 2)

Specify the sample size per treatment arm (assumed equal) and a matrix with allocation ratios in each period. The matrix with sample sizes per period is then computed based on these input parameters.

```{r, eval=FALSE}
# Example

head(data_sim_cont(n_total = NULL, num_arms = NULL, t_arm = NULL, 
                   n_arm = 200, alloc_ratios = , matrix(c(1,1,1,
                                                          1,1,NA,
                                                          NA,1,1), ncol = 3, byrow = T),
                   period_blocks = 2, delta = rep(0.25, 3), lambda = rep(0.15, 4), 
                   sigma = 1, trend = "linear"))
```

### Outputs

Define new parameter `full`. If `full=F`, then the output only contains response, treatment, period. Otherwise,then the output is all the dataframe. 



